---
title: "Precard report on Risk analysis"
author: "Tasos"
date: "25.05.2022"
output:
  html_document:
    toc: yes
    toc_depth: 3
    number_sections: true
    toc_float: 
      collapsed: false
    code_folding: hide
    theme: cerulean
---

```{r setup,echo=FALSE}
# Libraries
# knitrSet(lang='markdown', fig.path='png/', fig.align='left', w=6.5, h=4.5, cache=TRUE)
# If using blogdown: knitrSet(lang='blogdown')
knitr::opts_chunk$set(echo = FALSE)

library("dplyr")
library("survival")
library("knitr")
library('survivalAnalysis')
library("riskRegression")
library("rms")
library("survminer")

```

This report includes the analysis of the Type 1 diabetes nephropathy data, including survival models using 3 different packages for validation and reproducibility: RMS, RiskRegression and survival. All parameters were tested using backward variable selection and consecutive comparisons using appropriate statistics

```{r echo=FALSE}
html(contents(Precard_eGFRdrop_datasets$eGFR_drop50), maxlevels=10, levelType='table')
```

# 1. RMS package - Model

## Cox regression model

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="",  results='markup'}
# Development of the model using the fastbw() function of the RMS model to identify the optimal combination of variables 
rms.cox <- cph(Surv(T2E, EventCox) ~ Afrocaribbean_Ethnicity+Age_groups+ACR_G+SYSTOLIC+Hba1c, 
               data = Precard_eGFRdrop_datasets$eGFR_drop50, x=TRUE, y=TRUE, surv=TRUE, time.inc=1)
# Display model characteristics
rms.cox
# Fast Backward variable selection for backward elimination of factors. 
fastbw(rms.cox, rule = c("aic"))

```

## Hazard ratios estimates and 95% CI

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# Produce hazard ratios and 95% CIs for model parameters
cat("hazard ratio estimates")
pander(exp(coef(rms.cox))) # hazard ratio estimates)
cat("hazard ratio 95% CI")
pander(exp(confint(rms.cox))) # hazard ratio 95% CI
```

## Checking the Proportional Hazards Assumption

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# Proportionality of hazards check via plotting the beta(t) coefficients (line is horizontal) 
cox.zph(rms.cox)
survminer::ggcoxzph(cox.zph(rms.cox), var = c("Age_groups"), font.main = 12, ylab="Beta (t): Age (years)", xlab="Follow up (years)")
survminer::ggcoxzph(cox.zph(rms.cox), var = c("Afrocaribbean_Ethnicity"), font.main = 12, ylab="Beta (t): Ethnicity", xlab="Follow up (years)")
survminer::ggcoxzph(cox.zph(rms.cox), var = c("ACR_G"), font.main = 12, ylab="Beta (t): ACR levels", xlab="Follow up (years)")
survminer::ggcoxzph(cox.zph(rms.cox), var = c("SYSTOLIC"), font.main = 12, ylab="Beta (t): Systolic blood pressure", xlab="Follow up (years)")
survminer::ggcoxzph(cox.zph(rms.cox), var = c("Hba1c"), font.main = 12, ylab="Beta (t): Hba1c", xlab="Follow up (years)")
```

## Model evaluation

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# Use the anova.rms function for chiq test
anova(rms.cox, test = c( "Chisq"))

```

## Multicollinearity by Variance Inflation Factors

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

cat("Variance Inflation Factors")
pander(rms::vif(rms.cox))

```
## Validation of the model

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# Check difference in statistics such as R2, Somers’ d estimate etc, over cross-validated iterations

validate(rms.cox, method="crossvalidation", B=50, bw=FALSE, rule="aic", type="residual", sls=.05, 
         aics=0, force=NULL, estimates=TRUE, pr=FALSE, dxy=TRUE, tol=1e-9)

```

## Calibration plots for 1, 5 and 10 years

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# Create calibration plot to check if predicted survival probabilities from the model align with the observed survival probabilities
set.seed(43233)
cal.rms1<-rms::calibrate(rms.cox, cmethod=c('hare', 'KM'), method="boot", u=1, m=150, B=40, 
  bw=TRUE, rule="aic", type="residual", sls=0.05, aics=0, force=NULL,
  estimates=TRUE, pr=FALSE, what="observed-predicted", tol=1e-12, maxdim=5)
cal.rms5<-rms::calibrate(rms.cox5, cmethod=c('hare', 'KM'), method="boot", u=5, m=150, B=40, 
  bw=FALSE, rule="aic", type="residual", sls=0.05, aics=0, force=NULL,
  estimates=TRUE, pr=FALSE, what="observed-predicted", tol=1e-12, maxdim=5)
cal.rms10<-rms::calibrate(rms.cox10, cmethod=c('hare', 'KM'), method="boot", u=10, m=150, B=40, 
  bw=FALSE, rule="aic", type="residual", sls=0.05, aics=0, force=NULL,
  estimates=TRUE, pr=FALSE, what="observed-predicted", tol=1e-12, maxdim=5)

plot(cal.rms1, main="Predicted 1 year Survival", xlab="")
plot(cal.rms5, main="Predicted 5 years Survival", xlab="")
plot(cal.rms10, main="Predicted 10 years Survival", xlab="")

```

## Nomogram of the produced survival model

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup', fig.width=12, fig.asp = .82}

# “Packaging” data before the building of regression model and Nomogram by rms package. This is very important step for Nomogram building. 

dd=datadist(Precard_eGFRdrop_datasets$eGFR_drop50)
options(datadist="dd")
units(Precard_eGFRdrop_datasets$eGFR_drop50$T2E) <- "Year"

# Nomogram of survival rate will be built based on Cox regression. Set the survival object (surv1, surv2) based on survival function firstly.
surv1<- function(x)surv(1,lp = x) # defined time.inc 1 year survival probability 
surv2<- function(x)surv(5,lp = x) # defined time.inc 5 year survival probability

# Build survival Nomogram of Cox regression 
nom <- nomogram(rms.cox, fun = list(surv1, surv2), funlabel = c("1-year Survival Probability", "5-year Survival Probability"),
                abbrev = TRUE)
plot(nom, xfrac=0.3, cex.axis=.9, nint=20)

```

## PREDICTION with new instances

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# Use this prediction function to predict risk at different time points for "new" patient data including the same variables
predictSurvProb(rms.cox, newdata=test.data ,times=c(1, 5, 10, 14))

```

# 2. RiskRegression package

## Proportional hazards and fine-gray models

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# Cause specific Cox model
csc <- CSC(Hist(T2E, Event)~Afrocaribbean_Ethnicity+Age_groups+ACR_G+SYSTOLIC+Hba1c, data = Precard_eGFRdrop_datasets$eGFR_drop50, cause = "1")
# Print the model characteristics
csc

# Fine gray model
fgr <- FGR(Hist(T2E, Event)~Afrocaribbean_Ethnicity+Age_groups+ACR_G+SYSTOLIC+Hba1c, data = Precard_eGFRdrop_datasets$eGFR_drop50, cause = "1")
cat('Competeting Risks model with death as competing event:')
# print the model characteristics
fgr

```

## AUC and Brier plots of the models

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# (time-dependent) Brier score and the (time-dependent) AUC in the learning data using bootstrap
# cross-validation. The function optionally provides results for plotting (time-point specific)
##' ROC curves, for (time-point specific) calibration curves and for (time-point specific) retrospective boxplots.
score<-Score(list("Cause-specific Cox"=csc, "Fine-Gray"=fgr), formula = Hist(T2E,Event)~1,
              data=Precard_eGFRdrop_datasets$eGFR_drop50, times = seq(1,14,1), plots = "ROC", summary = "risks")
dev.new(width=5,height=4)
plotROC(score,times = 2)

```

## Prediction

### Prediction plot of the absolute risk of variables in a time frame

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

pred.csc<-predict(csc, newdata = test.data, time = 1:14, cause = 1)
autoplot(pred.csc)

```

### Absolute risk of ONE new patient at a specific time point

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# Absolute risk value

predictRisk(csc, newdata = test.data, time = 10, cause = 1)

```

### Absolute risk of multiple new patients at specific time point

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

pfit1 <- predict(csc, newdata = test.data, cause = 1, times = c(14), se = TRUE, keep.newdata = FALSE)
print(pfit1)

```

### Absolute risk alternative plot for 1 patient over 15 and 10 years

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

vec.times <- csc$eventTimes
pfit1band <- predict(csc, newdata = test.data, cause = 1, times = vec.times, se = TRUE, band = TRUE)

figure3 <- autoplot(pfit1band, band = TRUE, ci = TRUE)$plot
figure3 <- figure3 + xlab("Time (years)") + ylab("Absolute risk") + xlim(0,10)
print(figure3)

```


# 3. Survival package

## Summary of the cox model and anova comparison

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# CoxPH model for same variables
fit1<-coxph(Surv(T2E, EventCox) ~Afrocaribbean_Ethnicity+Age_groups+ACR_G+SYSTOLIC+Hba1c, 
               data = Precard_eGFRdrop_datasets$eGFR_drop50)

# Summary of the model and parameter characteristics
summary(fit1)

# Use anova() function to produce analyis of deviance table: 
anova(fit1)

```

## Plot of the survival model of the 15 year period

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

plot(survfit(fit1), ylab="Probability of Survival",
     xlab="Years", ylim = c(0.8, 1), col=c("red", "black", "black"))

```

## Proportional hazards assumption check + plot

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# Chisq test to check the proportionality of hazards.
cox.zph(fit1)

# plot time-dependent beta coefficient
plot(cox.zph(fit1))

#We’re looking for the smooth curve to be fairly level across the time horizon here, as opposed to substantially increasing or decreasing in level as time passes.

```


# End of code ----