---
title: "Precard Type 1 Tables for Publication"
author: "Tasos Mangelis"
date: "15.06.2022"
output: word_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library("dplyr")
library("survival")
library("knitr")
library('survivalAnalysis')
library("riskRegression")
library("rms")
library("survminer")
library("gtsummary")
library("pander")

```

\newpage
## Table 1 contains the basic patient characteristics following cleaning of the data and imputation as described in the methods section. 

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# use the gtsummary() function in the dplyr pipeline for nice tables
Precard_eGFRdrop_datasets$eGFR_drop50%>%
  select(Age, Diabetes_Dur, eGFR_bsl, SYSTOLIC, DIASTOLIC, Hba1c, ACR,
         Cholesterol, BMI, Weight, Ethnicity_groups, Afrocaribbean_Ethnicity, Gender)%>%
  tbl_summary(missing = "no", # Don;t include missing values
              # Choose desirable statistics to present
              statistic = list(all_continuous() ~ "{mean} ({sd}), {median} ({p25}, {p75})",
                     all_categorical() ~ "{n} ({p}%)"),
              # Change labels for best appearance
              label = list(Age ~	"Age (years)",
                           Diabetes_Dur~	"Duration of diabetes",
                           eGFR_bsl~	"eGFR baseline",
                           SYSTOLIC ~	"Systolic blood pressure",
                           DIASTOLIC~	"Diastolic blood pressure",
                           Hba1c	~	"HbA1c",
                           ACR ~	"ACR",
                           Cholesterol~	"Cholesterol",
                           BMI ~ "BMI", 
                           Afrocaribbean_Ethnicity~	"African Caribbean"	))%>%
  add_n() # add numbers to table


```

\newpage
## Table 2: Table with baseline patient caracteristics stratified by African-Caribbean Ethnicity

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Create table same as above
Precard_eGFRdrop_datasets$eGFR_drop50%>%
  select(Age, Diabetes_Dur, eGFR_bsl, SYSTOLIC, DIASTOLIC, Hba1c, ACR,
         Cholesterol, BMI, Weight, Ethnicity_groups, Afrocaribbean_Ethnicity, Gender)%>%
  # Additionally, use Ethnicity to divide data to two groups and apply all below to both
  tbl_summary(missing = "no", by = Afrocaribbean_Ethnicity, 
              statistic = list(all_continuous() ~ "{mean} ({sd}), {median} ({p25}, {p75})",
                     all_categorical() ~ "{n} ({p}%)"),
              label = list(Age ~	"Age (years)",
                           Diabetes_Dur~	"Duration of diabetes",
                           eGFR_bsl~	"eGFR baseline",
                           SYSTOLIC ~	"Systolic blood pressure",
                           DIASTOLIC~	"Diastolic blood pressure",
                           Hba1c	~	"HbA1c",
                           ACR ~	"ACR",
                           Cholesterol~	"Cholesterol",
                           BMI ~ "BMI", 
                           Afrocaribbean_Ethnicity~	"African Caribbean"	))%>%
  add_n()%>%
  add_p() # add statistical tests (Displayed at the footnote area of the table)

```

\newpage
## Table 2: Table with baseline patient caracteristics stratified by eGFR 50 percent drop


```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Create gtsummary table as before
Precard_eGFRdrop_datasets$eGFR_drop50%>%
  select(Age, Diabetes_Dur, eGFR_bsl, SYSTOLIC, DIASTOLIC, Hba1c, ACR,
         Cholesterol, BMI, Weight, Ethnicity_groups, Afrocaribbean_Ethnicity, Gender, EventCox)%>%
  mutate(EventCox=factor(EventCox, levels = c(1, 2), 
                         labels=c("progression", "cencored")))%>%
  tbl_summary(missing = "no", by = EventCox,
              statistic = list(all_continuous() ~ "{mean} ({sd}), {median} ({p25}, {p75})",
                     all_categorical() ~ "{n} ({p}%)"),
              label = list(Age ~	"Age (years)",
                           Diabetes_Dur~	"Duration of diabetes",
                           eGFR_bsl~	"eGFR baseline",
                           SYSTOLIC ~	"Systolic blood pressure",
                           DIASTOLIC~	"Diastolic blood pressure",
                           Hba1c	~	"HbA1c",
                           ACR ~	"ACR",
                           Cholesterol~	"Cholesterol",
                           BMI ~ "BMI", 
                           Afrocaribbean_Ethnicity~	"African Caribbean"	))%>%
  add_n()%>%
  add_p()

```

\newpage
## 50% drops Regression models

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='asis'}

# Use finalfit package for checking, but also because it creates better table outputs
# that allow all models to be presented at one table

explanatory = c("Age_groups", "Afrocaribbean_Ethnicity", "SYSTOLIC", "Hba1c","ACR_G") # Define variables
dependent_dss = "Surv(T2E,EventCox)" # Define events for Univariable and multivariable models
dependent_crr = "Surv(T2E,Event)" # Define survival objects for Compreting risks (Pre-event death)

Precard_eGFRdrop_datasets$eGFR_drop50%>%
  # Create summary table to include all results
  summary_factorlist(dependent_dss, explanatory, column = TRUE, fit_id = TRUE) %>%
    # CPH univariable models
  ff_merge(Precard_eGFRdrop_datasets$eGFR_drop50 %>% coxphuni(dependent_dss, explanatory) %>%
     fit2df(estimate_suffix = " (CPH univariable)")) %>%
    # CPH multivariable model
  ff_merge(Precard_eGFRdrop_datasets$eGFR_drop50 %>% coxphmulti(dependent_dss, explanatory) %>%
      fit2df(estimate_suffix = " (CPH multivariable)")) %>%
    # Fine - Gray competing risks regression
  ff_merge(Precard_eGFRdrop_datasets$eGFR_drop50 %>% crrmulti(dependent_crr, explanatory) %>%
    fit2df(estimate_suffix = " (Competing risks model)")) %>%
  select(-fit_id, -index) %>% 
  dependent_label(Precard_eGFRdrop_datasets$eGFR_drop50, "Survival") %>% 
  kable()

```

\newpage
## Forest plot for 50% drop model

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='asis'}
# Create a forest plot including the Hazard ratios and CIS for all variables includeed in the model

tiff("Figure2_DiabetesCare.tiff", res = 700, width = 6000, height = 4000, units = "px") # save as high quality for journal
Precard_eGFRdrop_datasets$eGFR_drop50%>%
  mutate(SYSTOLIC = ff_label(SYSTOLIC, "Systolic blood pressure"))%>%
  mutate(Age = ff_label(Age_groups, "Age groups (years)"))%>%
  mutate(ACR =ff_label(ACR_G, "ACR grade"))%>%
  mutate(Afrocaribbean_Ethnicity = factor(Afrocaribbean_Ethnicity, 
                              levels=c("1. Non African-Caribbean", "2. African-Caribbean"), 
                              labels=c("Non African-Caribbean", "African-Caribbean")))%>%
  hr_plot(dependent_dss, explanatory, remove_ref = T,
          dependent_label = NULL, suffix = NULL, prefix = NULL, 
          table_text_size=3.5, title_text_size=12, 
          plot_opts=list(xlab("HR, 95% CI"), theme(axis.title = element_text(size=12))))
dev.off()

```

\newpage
## Cumulative incidence curve for Afro-caribbean ethnicity

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

# Create a surv.fit object for ethnicity
fit.test<-survfit(Surv(T2E, EventCox) ~ Afrocaribbean_Ethnicity, data = Precard_eGFRdrop_datasets$eGFR_drop50)

# Create a plot object using ggsurvplot()
ggsurv_fig1<-ggsurvplot(fit.test,
          # plot specifics
          conf.int = F,
          linetype = "solid",
          ggtheme = theme_classic(),
          palette = c("#999999", "#E69F00"),
          # Legend specifics
          legend.labs=c("Non African-Caribbean", "African-Caribbean"),
          legend.title = "Ethnicity",
          legend.fontsize=12,
          fun = "cumhaz",
          # axis and labels specifics
          xlim=c(0, 13.9), ylim=c(0, 0.5), 
          font.x = c(14),
          pval = F,
          font.y = c(14),
          font.tickslab = c(14, "plain"), 
          xlab="Follow up (years)",
          ylab="Cumulative Hazard",
          risk.table = T,
          tables.theme = theme_cleantable(),
          tables.y.text = FALSE,
          break.time.by = 2, 
          censor.shape="")

# Further define label for being statistically significant and place between the figures
ggsurv_fig1$plot <- ggsurv_fig1$plot +
     annotate("text", x = 14.5, y = 0.24, label = "p<0.001", cex=3, col="black", vjust=0, hjust = 1.1, fontface=1.5)

# Save figure in good quality .tiff for publisher requirements
tiff("Figure1_DiabetesCare.tiff", res = 700, width = 5500, height = 4000, units = "px")
ggsurv_fig1
dev.off()
```


\newpage
## Cumulative incidence curve for Afro-caribbean ethnicity - 40%

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}
# produce similar plots for the other 2 models of 30 and 40% decrease
fit.test<-survfit(Surv(T2E, EventCox) ~ Afrocaribbean_Ethnicity, data = Precard_eGFRdrop_datasets$eGFR_drop40)
ggsurv40<-ggsurvplot(fit.test, 
          conf.int = F,
          linetype = "solid",
          ggtheme = theme_classic(),
          palette = c("#999999", "#E69F00"),
          legend.labs=c("Non African-Caribbean", "African-Caribbean"),
          fun = "cumhaz",
          xlim=c(0, 13.9), ylim=c(0, 0.5), 
          font.x = c(14),
          pval = T,
          font.y = c(14),
          font.tickslab = c(12, "plain"), 
          xlab="Follow Up (years)",
          ylab="Cumulative Hazard", 
          risk.table = T, 
          break.time.by = 2, 
          censor.shape="")

ggsurv40
```

\newpage
## Cumulative incidence curve for Afro-caribbean ethnicity - 30%

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

fit.test<-survfit(Surv(T2E, EventCox) ~ Afrocaribbean_Ethnicity, data = Precard_eGFRdrop_datasets$eGFR_drop30)
ggsurv30<-ggsurvplot(fit.test,
          conf.int = F,
          linetype = "solid",
          ggtheme = theme_classic(),
          palette = c("#999999", "#E69F00"),
          legend.labs=c("Non African-Caribbean", "African-Caribbean"),
          fun = "cumhaz",
          xlim=c(0, 13.9), ylim=c(0, 0.5), 
          font.x = c(14),
          pval = T,
          font.y = c(14),
          font.tickslab = c(12, "plain"), 
          xlab="Follow Up (years)",
          ylab="Cumulative Hazard", 
          risk.table = T, 
          break.time.by = 2, 
          censor.shape="")

ggsurv30
```

\newpage
## Additional info for paper

\newpage
### events per patient years for primary outcome (50% decrease)

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

Precard_eGFRdrop_datasets$eGFR_drop50%>%
  select(Event, T2E)%>%
  summarise(Count = sum(T2E)/1000, 
            Sum=sum(Event=="1"), 
            events.per.year=Sum/Count)%>%
  kable()
  
        
```

### events per patient years per ethnicity for primary

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

Precard_eGFRdrop_datasets$eGFR_drop50%>%
  select(Event, T2E, Afrocaribbean_Ethnicity)%>%
  group_by(Afrocaribbean_Ethnicity) %>%  
  mutate(Count = sum(T2E)/1000, 
         Sum_50 = sum(Event=="1"), 
         events.per.year= round((Sum_50/Count),2))%>%
  select(events.per.year, Afrocaribbean_Ethnicity)%>%
  tbl_summary(by = Afrocaribbean_Ethnicity)%>%
  add_p()

```




# End of code ----