---
title: "Precard Type 1 supplementary tables for publication"
author: "Tasos"
date: "15.06.2022"
output: word_document
---


```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("dplyr")
library("survival")
library("knitr")
library('survivalAnalysis')
library("riskRegression")
library("rms")
library("survminer")
library("gtsummary")
library("pander")
```


\newpage

# Tables for the secondary outcomes

\newpage
## Table 2: Table with baseline patient caracteristics stratified by eGFR 40 percent drop


```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Create gtsummary table as before
Precard_eGFRdrop_datasets$eGFR_drop40%>%
  select(Age, Diabetes_Dur, eGFR_bsl, SYSTOLIC, DIASTOLIC, Hba1c, ACR,
         Cholesterol, BMI, Weight, Ethnicity_groups, Afrocaribbean_Ethnicity, Gender, EventCox)%>%
  mutate(EventCox=factor(EventCox, levels = c(1, 2), 
                         labels=c("progression", "cencored")))%>%
  tbl_summary(missing = "no", by = EventCox,
              statistic = list(all_continuous() ~ "{mean} ({sd}), {median} ({p25}, {p75})",
                     all_categorical() ~ "{n} ({p}%)"),
              label = list(Age ~	"Age (years)",
                           Diabetes_Dur~	"Duration of diabetes",
                           eGFR_bsl~	"eGFR baseline",
                           SYSTOLIC ~	"Systolic blood pressure",
                           DIASTOLIC~	"Diastolic blood pressure",
                           Hba1c	~	"HbA1c",
                           ACR ~	"ACR",
                           Cholesterol~	"Cholesterol",
                           BMI ~ "BMI", 
                           Afrocaribbean_Ethnicity~	"African Caribbean"	))%>%
  add_n()%>%
  add_p()

```

\newpage
## Table 2: Table with baseline patient caracteristics stratified by eGFR 30 percent drop


```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Create gtsummary table as before
Precard_eGFRdrop_datasets$eGFR_drop30%>%
  select(Age, Diabetes_Dur, eGFR_bsl, SYSTOLIC, DIASTOLIC, Hba1c, ACR,
         Cholesterol, BMI, Weight, Ethnicity_groups, Afrocaribbean_Ethnicity, Gender, EventCox)%>%
  mutate(EventCox=factor(EventCox, levels = c(1, 2), 
                         labels=c("progression", "cencored")))%>%
  tbl_summary(missing = "no", by = EventCox,
              statistic = list(all_continuous() ~ "{mean} ({sd}), {median} ({p25}, {p75})",
                     all_categorical() ~ "{n} ({p}%)"),
              label = list(Age ~	"Age (years)",
                           Diabetes_Dur~	"Duration of diabetes",
                           eGFR_bsl~	"eGFR baseline",
                           SYSTOLIC ~	"Systolic blood pressure",
                           DIASTOLIC~	"Diastolic blood pressure",
                           Hba1c	~	"HbA1c",
                           ACR ~	"ACR",
                           Cholesterol~	"Cholesterol",
                           BMI ~ "BMI", 
                           Afrocaribbean_Ethnicity~	"African Caribbean"	))%>%
  add_n()%>%
  add_p()

```

\newpage
# COX MODELS - Tables for the Multivariate Model of the primary

\newpage
## Cox regression model

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="",  results='markup'}
dd=datadist(df_strat_finalfit)
options(datadist="dd")
units(df_strat_finalfit$T2E) <- "Year"

rms.cox <- cph(Surv(T2E, EventCoxF) ~ Age_groups+Afrocaribbean_Ethnicity+ACR_levels+Systolic_mmHg+
                 `Hba1c`, data = df_strat_finalfit, x=TRUE, y=TRUE, surv=TRUE, time.inc=1)
pander(rms.cox)
cat("hazard ratio estimates")
pander(exp(coef(rms.cox))) # hazard ratio estimates)
cat("hazard ratio 95% CI")
pander(exp(confint(rms.cox))) # hazard ratio 95% CI
```

\newpage
### Checking the Proportional Hazards Assumption

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

cox.zph(rms.cox)

survminer::ggcoxzph(cox.zph(rms.cox), var = c("Age_groups"), font.main = 12, ylab="Beta (t): Age (years)", xlab="Follow up (years)")
survminer::ggcoxzph(cox.zph(rms.cox), var = c("Afrocaribbean_Ethnicity"), font.main = 12, ylab="Beta (t): Ethnicity", xlab="Follow up (years)")
survminer::ggcoxzph(cox.zph(rms.cox), var = c("ACR_G"), font.main = 12, ylab="Beta (t): ACR levels", xlab="Follow up (years)")
survminer::ggcoxzph(cox.zph(rms.cox), var = c("SYSTOLIC"), font.main = 12, ylab="Beta (t): Systolic blood pressure", xlab="Follow up (years)")
survminer::ggcoxzph(cox.zph(rms.cox), var = c("Hba1c"), font.main = 12, ylab="Beta (t): Hba1c", xlab="Follow up (years)")



```

\newpage
### Multicollinearity by Variance Inflation Factors

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

cat("Variance Inflation Factors")
pander(rms::vif(rms.cox))

```


\newpage
## Model evaluation

\newpage
### Anova for model evaluation

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}

pander(anova(rms.cox))

```

\newpage
### Validation of the model for 1, 5 and 10 years by visual observation

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='markup'}
cat("Validation for the 1 year model")
rms::val
validate(rms.cox, method="boot", bw=FALSE, rule="aic", type="residual", sls=.05, 
         aics=0, force=NULL, estimates=TRUE, pr=FALSE, dxy=TRUE, tol=1e-9)

cat("Validation for the 5 year model")
validate(rms.cox5, method="boot", bw=FALSE, rule="aic", type="residual", sls=.05, 
         aics=0, force=NULL, estimates=TRUE, pr=FALSE, dxy=TRUE, tol=1e-9)

cat("Validation for the 10 year model")
validate(rms.cox10, method="boot", bw=FALSE, rule="aic", type="residual", sls=.05, 
         aics=0, force=NULL, estimates=TRUE, pr=FALSE, dxy=TRUE, tol=1e-9)
#The R2 statistic barely moves, and neither does the Somers’ d estimate, so at least in this simple model, the nominal summary statistics are likely to hold up pretty well in new data

```


\newpage
## 40% drops Regression models  

```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='asis'}
explanatory = c("Age_groups", "Afrocaribbean_Ethnicity", "SYSTOLIC", "Hba1c","ACR_G") # Define variables
dependent_dss = "Surv(T2E,EventCox)" # Define events for Univariable and multivariable models
dependent_crr = "Surv(T2E,Event)" # Define survival objects for Compreting risks (Pre-event death)

Precard_eGFRdrop_datasets$eGFR_drop40%>%
  # Create summary table to include all results
  summary_factorlist(dependent_dss, explanatory, column = TRUE, fit_id = TRUE) %>%
    # CPH univariable models
  ff_merge(Precard_eGFRdrop_datasets$eGFR_drop40 %>% coxphuni(dependent_dss, explanatory) %>%
     fit2df(estimate_suffix = " (CPH univariable)")) %>%
    # CPH multivariable model
  ff_merge(Precard_eGFRdrop_datasets$eGFR_drop40 %>% coxphmulti(dependent_dss, explanatory) %>%
      fit2df(estimate_suffix = " (CPH multivariable)")) %>%
    # Fine - Gray competing risks regression
  ff_merge(Precard_eGFRdrop_datasets$eGFR_drop40 %>% crrmulti(dependent_crr, explanatory) %>%
    fit2df(estimate_suffix = " (Competing risks model)")) %>%
  select(-fit_id, -index) %>% 
  dependent_label(Precard_eGFRdrop_datasets$eGFR_drop40, "Survival") %>% 
  kable()


# Create a forest plot including the Hazard ratios and CIS for all variables includeed in the model

tiff(paste0(Datasave_file_location,"Forestplot_40p.tiff"), 
     res = 700, width = 6000, height = 4000, units = "px") # save as high quality for journal
Precard_eGFRdrop_datasets$eGFR_drop40%>%
  mutate(SYSTOLIC = ff_label(SYSTOLIC, "Systolic blood pressure"))%>%
  mutate(Age = ff_label(Age_groups, "Age groups (years)"))%>%
  mutate(ACR =ff_label(ACR_G, "ACR grade"))%>%
  mutate(Afrocaribbean_Ethnicity = factor(Afrocaribbean_Ethnicity, 
                              levels=c("1. Non African-Caribbean", "2. African-Caribbean"), 
                              labels=c("Non African-Caribbean", "African-Caribbean")))%>%
  hr_plot(dependent_dss, explanatory, remove_ref = T,
          dependent_label = NULL, suffix = NULL, prefix = NULL, 
          table_text_size=3.5, title_text_size=12, 
          plot_opts=list(xlab("HR, 95% CI"), theme(axis.title = element_text(size=12))))
dev.off()
```

\newpage

## 30% drops Regression models
```{r echo=FALSE, message=FALSE, warning=FALSE, comment="", results='asis'}
explanatory = c("Age_groups", "Afrocaribbean_Ethnicity", "SYSTOLIC", "Hba1c","ACR_G") # Define variables
dependent_dss = "Surv(T2E,EventCox)" # Define events for Univariable and multivariable models
dependent_crr = "Surv(T2E,Event)" # Define survival objects for Compreting risks (Pre-event death)

Precard_eGFRdrop_datasets$eGFR_drop30%>%
  # Create summary table to include all results
  summary_factorlist(dependent_dss, explanatory, column = TRUE, fit_id = TRUE) %>%
    # CPH univariable models
  ff_merge(Precard_eGFRdrop_datasets$eGFR_drop30 %>% coxphuni(dependent_dss, explanatory) %>%
     fit2df(estimate_suffix = " (CPH univariable)")) %>%
    # CPH multivariable model
  ff_merge(Precard_eGFRdrop_datasets$eGFR_drop30 %>% coxphmulti(dependent_dss, explanatory) %>%
      fit2df(estimate_suffix = " (CPH multivariable)")) %>%
    # Fine - Gray competing risks regression
  ff_merge(Precard_eGFRdrop_datasets$eGFR_drop30 %>% crrmulti(dependent_crr, explanatory) %>%
    fit2df(estimate_suffix = " (Competing risks model)")) %>%
  select(-fit_id, -index) %>% 
  dependent_label(Precard_eGFRdrop_datasets$eGFR_drop30, "Survival") %>% 
  kable()


# Create a forest plot including the Hazard ratios and CIS for all variables includeed in the model

tiff(paste0(Datasave_file_location,"Forestplot_30p.tiff"), 
     res = 700, width = 6000, height = 4000, units = "px") # save as high quality for journal
Precard_eGFRdrop_datasets$eGFR_drop30%>%
  mutate(SYSTOLIC = ff_label(SYSTOLIC, "Systolic blood pressure"))%>%
  mutate(Age = ff_label(Age_groups, "Age groups (years)"))%>%
  mutate(ACR =ff_label(ACR_G, "ACR grade"))%>%
  mutate(Afrocaribbean_Ethnicity = factor(Afrocaribbean_Ethnicity, 
                              levels=c("1. Non African-Caribbean", "2. African-Caribbean"), 
                              labels=c("Non African-Caribbean", "African-Caribbean")))%>%
  hr_plot(dependent_dss, explanatory, remove_ref = T,
          dependent_label = NULL, suffix = NULL, prefix = NULL, 
          table_text_size=3.5, title_text_size=12, 
          plot_opts=list(xlab("HR, 95% CI"), theme(axis.title = element_text(size=12))))
dev.off()
```

# End of code ----